\titleNameLong{}

### 2015-12-07 FinalProject/Tex/paper/paper.tex
worked on architecture section

\section{Architectural Overview} \label{sec:architecture}
\subsection{Streaming Multiprocessors}
Modern unified GPUs consist of multiple streaming multiprocessors (SMs)
\begin{itemize}
  \item warp (NVIDIA)/wavefronts (OpenCL)
  \item similar to a SIMD processor
  \item SM has a warp pool to store context of all running threads
  \item HW warp scheduler dispatches and allocates resources
  \item large (much larger than a CPU) register file shared across all threads
    \item Maxwell architecture has 64K 32-bit registers for each SM
  \item kernel: large number of threads executing same piece of code
  \item multiple threads form a thread-block; same life-cycle, dispatched together
  \item threads within a thread-block (32 or 64) organized into warps
  \item warp scheduler schedules a ready warp each cycle, hiding latency of stalled warps
  \item CPU launches GPU kernels by dispatching launch commands. Kernel parameters are passed from GPU to CPU at launch time and stored in GPU global memory
  \item kernels from different streams are independent and can be run concurrently, but kernels within a stream are sequential
  \item see DTBL paper for good intro
  \item also see Locality Aware paper for internals of a SM
  \item cooperative thread array (CTA) - see Weak behavior paper
\end{itemize}
\subsection{NVIDIA Warps}
\begin{itemize}
  \item Multiple threads are grouped together in ``warps''
  \item Thread within a warp are mapped to a vector functional unit (SIMD)
  \item NVIDIA's Maxwell GPUs support up to 64 concurrent warps (2048 threads)
\end{itemize}

>>>>

\section{Architectural Overview} \label{sec:architecture}
\subsection{Streaming Multiprocessors}
\begin{itemize}
  \item Modern unified GPUs consist of multiple streaming multiprocessors (SMs)
  \item similar to a SIMD processor
  \item A SM contains
  \begin{itemize}
    \item a large register file shared across all threads. This register file
    typically contains thousands of registers (the NVIDIA Maxwell architecture
    has 64K 32-bit registers in each SM)
    \item multiple caches. Maxwell includes a unified L1/texture cache used as a
    coalescing buffer for memory accesses. The buffer gathers requested data and
    sends the entire group to a warp.
    \item shared memory (96KB per SM). Thread blocks are limited to 48KB each.
    Shared memory for a thread-block is only accessible by threads within the
    same thread-block.
    \item multiple warp schedulers. A warp scheduler dispatches and allocates
    resources. Maxwell has four per SM. FUs are assigned to a particular
    scheduler. Power of two CUDA cores per SM allows for simple partitioning
    among schedulers. One ready warp is scheduled per cycle, hiding the latency
    of stalled warps.
  \end{itemize}

\end{itemize}
\subsection{Kernels, Thread-blocks, and Warps}
\begin{itemize}
  \item A kernel is a massive number of threads executing same piece of code.
  Kernels from different streams are independent and can be run concurrently,
  but kernels within a stream are sequential.
  \item The CPU launches GPU kernels by dispatching launch commands. Kernel
  parameters are passed from the GPU to CPU at launch time and are stored in the
  GPU's global memory.
  \item Grids are what actually run on the GPU and are composed of multiple
  ``thread- blocks'', also called a cooperative thread array (CTA). Thread-
  blocks contain on the order of 512 threads. Each thread within a thread-block
  has the same life-cycle as all the others, meaning that they are dispatched to
  and swapped from a SM at the same time.
  \item Individual threads exist within a thread-block. These threads are
  grouped, usually around 32 at a time, into ``warps''. The warp is the
  fundamental unit of execution on GPUs. Warp is NVIDIA's term; AMD and OpenCL
  use the term ``wavefront''. Maxwell supports up to 64 warps resident per SM.
  This doesn't mean that all 64 are executing at the same time, but that the
  architecture can support that many warp contexts concurrently.
  \item The SM maintains a warp pool to store the context of all running
  threads. Warps are scheduled by the warp schedulers from this pool.
  \item Threads within a warp are mapped to a vector functional unit such that
  all threads execute the same instruction on different data, just a like a SIMD
  unit.
\end{itemize}
